---
title: 'Portfolio Positions Update'
author: "Team Greenleaf"
date: "2018-12-31"
output: html_document
description: "Our portfolio positions update"
---

```{r lib,message=FALSE,warning=FALSE,echo=FALSE}

library(tidyquant)
library(ggthemes)
library(dygraphs)
library(timetk)
library(tibbletime)
library(DT)


```


On October 3rd we published our stock portfolio positions. Q4 of 2018 saw large pullback in the markets. In hindsight it appears that October 2018 may have put in a top (at least until we see that level breached). Given the recent pullback and the uncertainty in the future, we have sold all our stock positions today and switched all our portfolios into bonds. We invested 100% our portfolios in Aggregate Bond Index ETF (symbol AGG)


### Performance for Portfolios and Market Q4 2018

```{r index-price,message=FALSE,warning=FALSE,echo=FALSE}

indx_price <- read_csv("E:/Documents/greenleaf_web_data/Dec/Dec31_Position_update/index_price_data.csv")

```



```{r pos-data, message=FALSE,warning=FALSE,echo=FALSE}

spx_pos <- read_csv("E:/Documents/greenleaf_web_data/Dec/Dec31_Position_update/spx_pos.csv")
dow_pos <- read_csv("E:/Documents/greenleaf_web_data/Dec/Dec31_Position_update/dow_pos.csv")
rus_pos <- read_csv("E:/Documents/greenleaf_web_data/Dec/Dec31_Position_update/rus_pos.csv")


```


```{r pos-returns, message=FALSE,warning=FALSE,echo=FALSE}
port_pnl <- tibble(port = c("Dow Jones Portfolio", "S&P 500 Portfolio", "Russell Portfolio"),
       value = c(sum(dow_pos$value), sum(spx_pos$value), sum(rus_pos$value)),
       PnL = c(sum(dow_pos$PnL), sum(spx_pos$PnL), sum(rus_pos$PnL)))  

port_pnl <- port_pnl %>%
  mutate(returns = PnL/value)

index_ret <- indx_price %>%
  group_by(symbol) %>%
  tq_transmute(select = price,
               mutate_fun = periodReturn,
               period = "quarterly",
               col_rename = "returns") %>%
  spread(symbol, value = returns) %>%
  rename(`Dow Jones Index` = DIA,
         `S&P 500 Index` = SPY,
         `Russell 2000 Index` = IWM) %>%
  gather(`Dow Jones Index`:`S&P 500 Index`,key = Index, value = returns) 


port_pnl %>%
  rename(Name = port) %>%
  select(Name,
         returns) %>%
  bind_rows(index_ret %>%
              select(-date) %>%
            rename(Name = Index)) %>%
  datatable(caption = "Our Portfolio returns vs the Index returns since October 3, 2018", rownames = FALSE,
            colnames = ("Returns" = 'returns')) %>%
  formatPercentage(columns = "returns", digits = 2)


```

<br/>

We can see that the indexes and the portfolios had a big draw downs, but our portfolios underperformed their respective benchmarks.

Next we will see the individual performance for all the stock holdings in each of the portfolios.

<br/>
<br/>

```{r pos-table,message=FALSE,warning=FALSE,echo=FALSE}

dow_pos %>%
  select(-no) %>%
  rename(Symbol = symbol) %>%
  mutate(`PnL%` = PnL/value) %>%
  datatable(rownames = FALSE, caption = "Dow Jones Portfolio Positions and Performance since October 3, 2018",
            colnames = c("Symbol", "Name", "Trade Price", "Quantity",
                         "Investment Value", "Sell Price", "PnL", "PnL %")) %>%
  formatCurrency(currency = "$", digits =  2, columns = "PnL") %>%
  formatPercentage(columns = "PnL%")
  
  

```

<br/><br/>


```{r spx-table,message=FALSE,warning=FALSE,echo=FALSE}

spx_pos %>%
  select(-no) %>%
  rename(Symbol = symbol) %>%
  mutate(`PnL%` = PnL/value)  %>%
  datatable(rownames = FALSE, caption = "S&P 500 Portfolio Positions and Performance since October 3, 2018",
            colnames = c("Symbol", "Name", "Trade Price", "Quantity",
                         "Investment Value", "Sell Price", "PnL", "PnL %")) %>%
  formatCurrency(currency = "$", digits =  2, columns = "PnL") %>%
  formatPercentage(columns = "PnL%")

```

<br/>
<br/>


```{r rus-pos-table, message=FALSE,warning=FALSE,echo=FALSE}

rus_pos %>%
  select(-no) %>%
  rename(Symbol = symbol) %>%
  mutate(`PnL%` = PnL/value)  %>%
  datatable(rownames = FALSE, caption = "Russell Portfolio Positions and Performance since October 3, 2018",
            colnames = c("Symbol", "Name", "Trade Price", "Quantity",
                         "Investment Value", "Sell Price", "PnL", "PnL %")) %>%
  formatCurrency(currency = "$", digits =  2, columns = "PnL") %>%
  formatPercentage(columns = "PnL%")


```

<br/><br/>

As mentioned above we have sold all our positions and are now invested in Bonds. You can see the current ending value of our portfolios below.

```{r edn-val, message=FALSE,warning=FALSE,echo=FALSE}

port_pnl %>%
  mutate(end = PnL + value) %>%
  select(port, value, PnL, end) %>%
  datatable(colnames = c("Portfolio", "Portfolio Value on Oct 3rd", "PnL on Dec 31st", "Ending Value on Dec 31st"), rownames = FALSE)

end_val <- port_pnl %>%
  mutate(end = PnL + value) %>%
  select(port, value, PnL, end)



```

Current Portfolio is 100% invested in the Aggregate Bond index ETF which has the symbol AGG. You can see the value and quantity below.

```{r current-val, message=FALSE,warning=FALSE,echo=FALSE}

end_val %>%
  select(port,end) %>%
  mutate(`AGG price` = 106.40) %>%
  mutate(qty = end/`AGG price`) %>%
  mutate(qty = round(qty,-0.5)) %>%
  select(port, `AGG price`, qty, end) %>%
  datatable(caption = 'Current portfolios as of Dec 31st, 2018\n 100% invested in AGG', rownames = FALSE,
            colnames = c("Portfolios", "AGG price", "Quantity", "Market Value"))


```

We believe that the stock markets could see further draw downs going into 2019, and we believe its time to go defensive. We will change our opinion if/when we see positive data to support our thesis.

Wish you a Happy and prosperous New Year



